---
title: "Class Notes - Day 4"
author: "Charlie Hall"
date: "10/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(GGally)
library(ggcorrplot)
library(knitr)
library(hexbin)
library(descr)
```

### Sampling Distribution of a Statistic

**Sampling Distribution of the Sample Mean** - most common expression of sampling dist.

- big population (ex. $\mu = 82.5$)
- random samples (fixed sample size, $\overline{x}_{1} = 79.6$, $\overline{x}_{2} = 81$, etc.)
- sampling distribution of sample means ($79.6, 81.0, 85.2, 80.5, 79.8, 88.3$)
- avg. mean of sample means ($82.4$)

can also have sampling distribution of sample variance, sample median, so on.

**Sampling Dist**: frequency distribution or theoretical density curve of a sample stat over many samples with fixed sample size $n$.

**Central Limit Theorem (CLT)**: the tendency of the sampling dist. of the sample mean ($\overline{X}$) to take on a normal dist. as sample size rises.

**Standard Error**: variability (standard dev.) of a sample statistic over many samples 

- *standard deviation of the sampling distribution*

```{r}
library(nycflights13)
(dist_data <- flights %>% select(distance) %>%
    filter(!is.na(distance)))
```

```{r}
samp_mean_05 <- rep(0,1000)
for( i in 1:1000 ){
  samp_mean_05[i] <- mean( sample(dist_data$distance, 5 ))
}
hist(samp_mean_05, breaks = 15)

samp_mean_tib <- tibble(samp_mean_05)
ggplot(samp_mean_tib, aes(samp_mean_05)) +
  geom_histogram(bins = 15, color = 'black')
```

Sampling with size 25 & 64
```{r}
samp_mean_25 <- rep(0, 1000)
for(i in 1:1000){
  samp_mean_25[i] <- mean( sample(dist_data$distance, 25 ) )
}

samp_mn_25 <- tibble(samp_mean_25)

ggplot(samp_mn_25, aes(samp_mean_25)) +
  geom_histogram(bins = 15, color = 'black')

samp_mean_64 <- rep(0, 1000)
for(i in 1:1000){
  samp_mean_64[i] <- mean( sample(dist_data$distance, 64 ) )
}

samp_mn_64 <- tibble(samp_mean_64)

ggplot(samp_mn_64, aes(samp_mean_64)) +
  geom_histogram(bins = 15, color = 'black')

ggplot(dist_data) +
  geom_histogram(aes(distance), bins = 15)
```

**CLT** is not as important as it was (esp. in data sci).

- if population is normal, then sampling dist is normal
- if pop. is unknown/skewed, then sampling dist is close to normal when $n>25$.

**Standard Error**: $\frac{s}{\sqrt{n}}$
```{r}
single_sample_64 <- sample( dist_data$distance, 64 )
SE <- sd( single_sample_64 ) / sqrt( length( single_sample_64 ) ) 
SE
```

**Normal Distribution**: very common in nature (e.g. measures on biological specimens, standardized tests, the sampling distribution of the sample means, etc.)

Empirical rule: 1 standard dev. away from center covers about 68%. 2 sd's away covers about 95%. 3 sd's away covers 99.7%

**Standard Normal Dist**: normal with $\mu = 0$ and $\sigma = 1$.

$N(0,1)$

Z-score (converts to standard normal dist. so we can compare on a fixed scale):
$$z = \frac{x-\mu}{\sigma}$$

**Example**:

John's ACT score is 29.5

Jane's SAT score is 1431

ACT ~ N($\mu$ 20.8, $\sigma$ 5.8)

SAT ~ N($\mu$ 1051, $\sigma$ 211)

```{r}
(SATN <- (29.5-20.8) / 5.8)

(ACTN <- (1431 - 1051) / 211)
```

**QQ-Plot**: visually determine how close a sample is to a specified distribution - many cases, the normal dist.
```{r}
binom_sample <- rbinom( 100, 2500, 1/6)
qqnorm( binom_sample )
max(binom_sample)
min(binom_sample)
```