---
title: "Class Notes - Day 3"
author: "Charlie Hall"
date: "10/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(GGally)
library(ggcorrplot)
library(knitr)
library(hexbin)
library(descr)
```

### Correlation

**Covariance**: a measure of the joint variability of two numerical variables

**Population Covariance**: for discrete and continuous distributions, respectively

*no function for the population covariance*
```{r}
cov(iris$Sepal.Length[iris$Species=="setosa"], iris$Sepal.Width[iris$Species=="setosa"])
```

- Cannot compare the strength of association between two sets of data with their covariances. (.099 does not really mean anything other than that it's positive)
- magnitudes or units!
- Thus, only signs are important in comparing the strength of association between two sets of data, unless they save the same sets of units. 

**Correlation**: a measure of the joint variability of two numerical variables when divided by standard deviations (unitless, -1 <= $\rho$ <= 1)

*no function for the population correlation*
```{r}
cor( iris$Sepal.Length[iris$Species == 'setosa'], iris$Sepal.Width[iris$Species == 'setosa'])
```

```{r}
iris %>% ggplot( aes(x = Sepal.Length, y = Sepal.Width, color = Species) ) +
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE)
```

**Matrix of Scatter Plots**: if we have more than 2 numerical variables
```{r, message = FALSE}
mtcars %>% ggpairs()
```

```{r}
corr_tmp <- cor( mtcars )
corr_tmp %>% ggcorrplot( method = 'circle', hc.order = TRUE,
                         type = 'lower', outline.col = 'white')
```

```{r}
corr_tmp
```

**Linear** correlation!

- correlation zero doesn't mean no association
  - non-linear association may exist
- it measures linear correlation only

### Exploring 2 or more Variables

Tax-assessed values for residential properties in King County, Washington:
```{r, message - FALSE}
kc_tax <- read_csv('~/Downloads/kc_tax.csv')
kable( kc_tax[1:5, ])
```

Exclude very expensive homes, very small ones, or very large ones:
```{r}
kc_tax0 <- subset(kc_tax, TaxAssessedValue < 750000 & SqFtTotLiving > 100 & SqFtTotLiving < 3500)
nrow(kc_tax0)
```

For a very large dataset, we may use **Contour Plot**:
```{r}
ggplot(kc_tax0, aes(SqFtTotLiving, TaxAssessedValue)) +
  theme_bw() +
  geom_point(alpha = .1) +
  geom_density2d(color = 'white') +
  labs(x = 'Finished Square Feet', y = 'Tax-Assessed Value')
```

For a very large dataset, we may use **Hexagonal Binning Plot**:
```{r}
ggplot(kc_tax0, aes(SqFtTotLiving, TaxAssessedValue)) +
  stat_binhex(color = 'white') + theme_bw() +
  scale_fill_gradient(low = 'white', high = 'black') +
  labs( x = 'Finished Square Feet', y = 'Tax-Assessed Value' )
```

Grade of a personal loan vs. the outcome of that loan
```{r}
lc_loans <- read_csv("~/Downloads/lc_loans.csv")
head(lc_loans)

unique(lc_loans$status); unique(lc_loans$grade)
```

For two *categorical* variables, we use **Contingency Table**:
```{r}
CrossTable(lc_loans$grade, lc_loans$status,
           prop.c = FALSE, prop.chisq = FALSE, prop.t = FALSE)
```

For categorical vs. numerical, we use **Boxplots**:
```{r}
diamonds %>% ggplot( aes(x = cut, y = price)) +
  geom_boxplot()
```

For categorical vs. numerical, we may use enhanced boxplot with density estimates (**Violin Plot**):
```{r}
diamonds %>% ggplot( aes( cut, price )) +
  geom_violin() 
```

`facet_wrap` is useful for visualizing multiple variables:
```{r}
ggplot(subset(kc_tax0, ZipCode %in% c(98188,98105,98108,98126)),
       aes(SqFtTotLiving,TaxAssessedValue)) +
  stat_binhex(color ='white') +
  theme_bw() +
  scale_fill_gradient(low = 'white', high = 'blue') +
  labs(x = 'Finished Sq Feet', y = 'Tax-Assessed Value') +
  facet_wrap('ZipCode')
```

### Random Sampling and Sampling Bias

Population vs. Sample

- Population - normally theoretical, difficult to investigate due to limited resources
- Sample - we typically calculate statistics on a sample, using results to make reasonable inference upon population characteristics

Bias vs. Variance

- unbiased estimator - on average, estimate produces same value with population parameter
- trade-off between bias and variance

**Random Sampling** to avoid bias - crucial

- simple random sampling - can be difficult when it comes to procedure, resources
- stratified random sampling - randomly select from each strata
  - divide population into strata, run simple random sampling for each group 

Random sampling is *very* important even in the era of Big Data

- avoid bias!
- data quality is more important than data quantity

- Population Mean $\mu$
- Population Standard Deviation $\sigma$
- Sample Mean $\overline{x}$
- Sample Standard Deviation $s$

Why is $\overline{x}$ an unbiased estimator of the population mean $\mu$?

- assume you have random sampling
- on average, sample means will cancel out and center around pop. mean

### Selection Bias

**Selection Bias**: resulting from the way in which observations are selected (non-random sampling)

**Data Snooping**: Extensive hunting through data in search of something interesting.

- "if you torture your data long enough, sooner or later, it will confess"
- cherry-picking research questions after the fact.

**Vast Search Effect**: statistically significant observation occurs by chance because of sheer size of dataset.

- 1/1024 chance that 10 coin tosses will land heads just by chance
- ask 20,000 people to toss a coin 10 times

*be careful about selection bias, data snooping, vast search effect / don't do these!*

* more like an art than science when it comes to many real-world applications. "3 types of lies: lies, damned lies, and statistics" - people can use statistics to twist information if not cautious re: potential fallacies

**Regression to the Mean**: 'going back' to the mean

- rookies of the year, sophomore slump
- extreme observations tend to be followed by more central ones
- Francis Galton on genetic tendencies: extremely tall men tend to have shorter sons

**Linear Regression**: linear relationship is estimated between predictor variables and an outcome variable