---
title: "Class Notes - Day 5"
author: "Charlie Hall"
date: "10/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

### Review - Sampling Dist of a Statistic

In reality we only have *one* sample. 

- Works for the sampling dist of the sample mean
  - either normal pop or big n (n>25) required
- Sampling dist of the sample *median*:
  - if pop. is symmetric, then similar to that of sample mean
  - need to know the population density to figure out
- Sampling dist of the sample variance:
  - asymptotically follows a scaled $\chi^{2}$ dist. if population is normal or $n$ is big.
  
### Bootstrap Distribution

Recall that we tried to figure out how sample stat ($\overline{X},Median(X),S^{2}$, etc.) behaves by using the sampling distribution.

With the development of computer simluations, we can achieve the same goal easily by using the bootstrap distribution (resampling *from* sample).

#### How-To
In real life, we typically only get **one** sample. 
How can we find sample distribution? **Resample**

Resample multiple times to obtain "bootstrap samples"

**Bootstrap dist. has same spread as sample dist.**

#### Summary

Extremely convenient for us to construct sampling distribution by using the bootstrap dist.

- the shape is the same while its center may be slightly off
- and we know the true sampling dist. is centered exactly at the population parameter (e.g. $\mu$)

This works nicely, esp. when $n$ is small or population dist. is unknown

**Condition**: our sample must be a representative sample in comparison with the population (must accurately reflect shape of the pop.). I.e., use *random sampling*. 

Example:
```{r}
library(nycflights13)
( dist_data <- flights %>% select(distance) %>%
    filter( !is.na(distance) ) )
```

Sampling Distribution of Sample Mean with $n = 25$:
```{r}
set.seed(25); samp_mean_25 <- rep(0,10000)
for(i in 1:10000 ){
  samp_mean_25[i] <- mean(sample(dist_data$distance, 25, replace = TRUE))
}; hist( samp_mean_25, breaks = 15 ); sd(samp_mean_25)
```

Bootstrap Distribution:
1. Take only one sample
2. resample from this to create the bootstrap dist.
```{r, eval = FALSE}
set.seed(250); one_sample <- sample( dist_data$distance, 25, replace = TRUE )
samp_mean_250 <- rep(0,10000)
for(i in 1:10000){
  samp_mean_250[i] <- mean(sample(one_sample, 25, replace = TRUE))
  hist( samp_mean_250, breaks = 15); sd(samp_mean_250)
}
```

```{r}
library(boot)
one_sample <- sample( dist_data$distance, 25, replace = TRUE )
samplemean <- function(x, d) {
  return(mean(x[d]))
}
(boot_obj <- boot(one_sample, samplemean, R = 10000))
```

Repeated Resampling (bootstrapping) is computationally intensive and was not feasible before computers were widely available

**Bootstrapping implies sampling with replacement**

Bootstrap is very practical without extensive study of mathematical approximation to sampling distributions

### Confidence Interval

**Confidence Interval**: typical way to present estimates as an interval range for a population parameter ($\mu, \sigma^{2}$, etc.).

If the sampling dist. is hard to construct, then the bootstrap resampling is an effective way to construct confidence intervals.

**CI from the sampling dist**

- Assume only one sample with $n = 25$
- With this we know that sampling dist. of sample mean will be approximately normal with mean $\mu$ and standard error $\sigma /\sqrt{n}$
- This single sample produces $\overline{x}_{1}$ and $s$ as follows:
```{r}
mean(one_sample); sd(one_sample)
```
- since we do not know $\mu$ and $\sigma$, we assume the sampling dist will follow approximately the normal dist with mean $\overline{x}_{1}$ and st. error $s/\sqrt{25}$
- recall that 2 std. errors from center covers 95%
- we are 95% confident that the true $\mu$ is in the interval shown below:
```{r}
(CI_1 <- c(mean(one_sample) - 2* sd(one_sample)/sqrt(25), mean(one_sample) + 2* sd(one_sample)/sqrt(25)))
```
- we can also create as follows, but not feasible if we only have one sample
- however, we *can* construct the bootstrap dist. with only one sample! Thus, the bootstrap CI is as follows:
```{r}
(CI_2 <- c(quantile(samp_mean_25, .025, quantile(samp_mean_25, .975))))
```

**Standard Deviation**: use when evaluating spread of dataset

**Standard Error**: use when determining precision of sample mean or variability of mean distribution

### Various Distributions

- most dist. are *not* normal
- **Skewedness**: when one tail is longer than the other
  - right skewed - right tail is longer
  - left skewed - left tail is longer
- $t$-distribution: if we use $s$ instead of $\sigma$ for the sampling dist, then it gets slightly heavier tails than standard normal in reality. Thicker tails than a $z$-distribution
   - dt(), pt(), qt(), rt()
   - only **one** parameter: df (degrees of freedom) - this scales the thickness of the tails
