---
title: "Class Notes - Day 1"
author: "Charlie Hall"
date: "10/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# 10/6/20 day 1 class notes
library(tidyverse)
library(knitr)
(state <- read_csv("~/Downloads/state.csv", col_names = TRUE))

kable( state[1:5,] , caption = "Murder Rate by States (per 100,000")
```

### Mean: avg with equal weight
$$\overline{X} = \frac{ \sum_{i = 1}^n x_i }{n}$$

$\mu$ is used to speak of pop. mean. We use $\overline{x}$ for sample mean.

$\sigma$ = pop. standard dev.
$s$ = sample st. dev.

*We tend to use Greek letters when ref. to population

```{r}
mean( state$Population )
mean( state[['Population']])
```

```{r}
totalPop <- sum( state$Population )
totalPop/50
```

### Weighted Mean: avg with different weights
$$\overline{X}_w = \frac{ \sum_{i = 1}^n w_i x_i }{ \sum_{i = 1}^n w_i}$$

```{r}
mean( state$Murder.Rate )
weighted.mean( state$Murder.Rate, w = state$Population )
```

### Estimates of Location (Mostly Center)
**Median**: most typical value. 50% of obs above and 50% of obs below
$$x_1, x_2, x_3, \cdots , x_{n-2}, x_{n-1}, x_n$$

If n is odd,
$$\text{Median}(x) = x_{(n+1)/2}$$

If n is even,
$$\text{Median}(x) = \frac{ x_{n/2} + x_{n/2+1} }{2}$$

```{r message = FALSE}
median( state$Murder.Rate )

library( matrixStats )
weightedMedian( state$Murder.Rate, w = state$Population )
```

**When mean and median are very similar or the same, your distribution is symmetric.**

```{r}
state %>% ggplot() +
  geom_histogram(aes( Murder.Rate ), bins = 15)
state %>% arrange(desc(Murder.Rate))
```

**Percentile**: data value with P percent of obs. lies below.
```{r}
quantile( state$Murder.Rate, c( 0.1,0.25,0.5,0.98 ) )
```

Median is robust (*not sensitive to extreme values*). Mean is very susceptible to extreme values.

Median shows a better "typical" value. 

**Trimmed Mean**: avg without p extreme values on both sides.
$$\overline{X}_{trm} = \frac{ \sum_{i = p + 1}^{n-p } x_i }{n-2p}$$

```{r}
mean( state$Murder.Rate, trim = 2/50)
```

**Outliers**: data values which are very different from most of the data