---
title: "Class Notes - Day 2"
author: "Charlie Hall"
date: "10/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Definition of Probability
**Subj. Prop**: educated guess

- economist predicts a 10% chance that s&p500 index will go up next yr

**Exp. Prob**: empirical, uses the outcomes obtained by repeatedly performing experiment

- collect data, record, evaluate data to determine probability of event

**Classical Prob**: theoretical, can be calculated when all outcomes are equally likely

- prob of selecting king of diamonds from deck of standard playing cards is 1/52

**Probability is always between 0 and 1**

### Estimates of Variability (spread)
```{r}
library(tidyverse); library(ggplot2)
head(iris)
```

```{r}
iris %>% ggplot( aes(x = Sepal.Length)) +
  geom_histogram(bins = 15, fill = 'red', alpha = .3)
```

```{r}
iris %>% ggplot(aes(x = Sepal.Length)) +
  geom_histogram(bins = 15, fill = 'red', alpha = .3) +
  facet_wrap(~ Species)
```

**Deviations**: (errors, residuals) difference between each observed value & estimate of location ($\overline{x}$ or $\mu$). 

Formula = ($x_{i}-\overline{x}$)

```{r}
( m_S.L <- mean(iris$Sepal.Length))
```

```{r}
iris %>% select(Sepal.Length) %>%
  mutate(Deviation = Sepal.Length - m_S.L)

sum(iris$Deviation)
```

**Variance**: truly useful numerical measure of variability. Mean squared error. Sum of squared deviations from the mean divided by ($n-1$) or $N$. 

Population Variance - divide by sample size

Sample Variance - divide by sample size minus 1

Notice all Greek symbols in stats are fixed (though unknown to us)

Why $(n-1)$ in $s^2$:

- sample variance will vary sample by sample
- $s^2$: one single observed sample var.
- $S^2$: random variable to represent all possible sample var. with fixed $n$. 
- unbiased estimate: we want to make sure the expected value of this random variable is the same as $\sigma^2$.

$$E(S^2) = \sigma^2$$

```{r}
var(iris$Sepal.Length)
```

```{r}
library( sjstats )
var_pop(iris$Sepal.Length)
```

**Standard Deviation**: sqrt(s^2) = s

Variance and std. deviation are, by definition, *always positive values*. 

```{r}
sd( iris$Sepal.Length )

sd_pop( iris$Sepal.Length )
```

**Mean Absolute Deviation from the Mean**: Manhattan Norm, L1 Norm. mean of the absolute values of the deviations from the mean

**Median Absolute Deviation from the Median**: (more popular mad). the median of absolute values of the deviations from the median

- Median-MAD is robust to extreme values
- Mean-MAD is sensitive to extreme values

**Range**: max - min

**IQR**: Inter-Quartile Range. Q3 - Q1. Very robust measure for spread. Where the middle 50% of the dataset falls

**5 Number Summary**: min, Q1, Median, Q3, max

```{r}
mad( iris$Sepal.Length )
IQR( iris$Sepal.Length )
fivenum( iris$Sepal.Length )
```

### Exploring the Data Description

**Box Plot** (Box and Whiskers)

- use 5 number summary,
- 1.5 times IQR can be used to determine possible outliers
```{r}
boxplot( iris$Sepal.Length, horizontal = TRUE)

iris %>% ggplot() +
  geom_boxplot(aes(x = Sepal.Length))

iris %>% ggplot() +
  geom_boxplot(aes(x = Sepal.Length, y = Species))
```

```{r}
state <- read_csv("~/Downloads/state.csv", col_names = TRUE)
boxplot( state$Population, horizontal = TRUE)

state %>% ggplot() +
  geom_boxplot( aes(x = Population))
```

```{r}
breaks <- seq( from = min( state$Population ),
               to = max( state$Population ), length = 11)

pop_freq <- cut(state$ Population, breaks = breaks, right = TRUE, include.lowest = TRUE)

table(pop_freq)
```

**Histogram**: plot by freq. table
```{r}
hist(state$Population, breaks = breaks)

state %>% ggplot(aes(x = Population)) +
  geom_histogram( fill = 'red', alpha = .3, breaks = breaks)
```

**Density Plot**: smoothed version of the histogram. Many options possible for kernel density function
```{r}
state %>% ggplot( aes( x = Population )) +
  geom_histogram( aes(y =..density..),
                  fill = 'red', alpha = .3, bins = 15) +
  geom_density(color = 'blue' , size = 2)
```

### Exploring Categorical Data
**Mode**: most common value or category in a dataset (can also mean number of 'peaks' in histogram, etc. when applied to numerical variable)

**Expected Value**: mean value of a randomly varying variable with assigned probabilities

**Bar Chart**: Equivalent to histogram for categorical variables
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

**Pie Chart**: not a good idea, most of the time. 
```{r}
tmp <- diamonds %>% group_by( cut ) %>% summarize(n = n() )
pie(tmp$n)

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut)) +
  coord_polar()
```
